>

### 2019-09-28 04_estimation.Rmd
Rewrite econometrics to consider SDEM

---
title: "04_estimation"
author: "Greg Macfarlane"
date: "6/17/2019"
output: html_document
---

@@

```{r models}
write_rds(pa_models, "data/pa_models.rds")
write_rds(obesity_models, "data/obesity_models.rds")
```

>>>>

---
title: "Model "
author: "Greg Macfarlane"
date: "6/17/2019"
output: html_document
---

@@

```{r models}
write_rds(slx_vs_sdem, "data/slx_vs_sdem.rds")
write_rds(pa_models, "data/pa_models.rds")
write_rds(obesity_models, "data/obesity_models.rds")
```

### 2019-09-28 manuscript/manuscript.Rmd
Rewrite econometrics to consider SDEM

We predict the rate of physical activity ($y$) in a census tract as a
function of the tract's sociodemographic characteristics and choice-based
accessibility to parks with the following  specification:
\begin{equation}\label{eq:themodel}
 y_i = f(\vect{x}_i, C_i, g(W, \vect{y}, \vect{X}))
\end{equation}
where $\vect{x}_i$ is a vector of tract-level sociodemographic attributes,
$C_i$ is a choice-based accessibility measure for the tract computed from
Equation \ref{eq:cs} and $g(W, \vect{y}, X)$ represents the impact of spatial
neighborhood effects on the outcome.

Using tracts as the unit of observation in the model allows us to consider the
availability of parks at a sub-metropolitan level, but it also introduces the problem
of spatial correlation of unobserved errors and spatial dependence in the data
generating process [@Cliff1970; @Anselin1980]. Spatial correlation occurs
when spatially-distributed unobservable or missing variables (school quality,
neighborhood prestige, etc.) influence the modeled outcomes. Spatial dependence
occurs when the outcome *depends* on the outcome in neighboring areas; for
example, seeing neighbors exercising may encourage exercise. The relaxed spatial
Durbin model (SDM) proposed by @Burridge1981 controls for both processes:
\begin{equation}\label{eq:sdm}
  \vect{y}= \rho W\vect{y} + X\vect{\beta} + WX\vect{\gamma} + \vect{\epsilon}
\end{equation}
where $\rho$ is the estimated spatial correlation between the outcome variables
$y_i, y_j, \ldots  \in \vect{y}$, $W$ is a matrix specifying the spatial
relationship between observations $i$ and $j$ in the dataset,
$\vect{\beta}$ is a vector of estimable coefficients relating the attributes $X$
of an observation to its outcome, and $\vect{\gamma}$ is a vector of
estimated coefficients relating the attributes of an observation's *neighbors*
to its outcome. If only correlation and not dependence exists in the data
generating process, it may  be more efficient to estimate a restricted
spatial autoregressive model and therefore
save degrees of freedom in the estimation. @tra_autoregression_paper
illustrate a decision algorithm originally proposed by @Florax2003 which uses
a likelihood ratio test of the SDM against the restricted spatial error model (SEM):
\begin{align}
{\boldsymbol y} &= X{\boldsymbol \beta} + {\boldsymbol u},  {\boldsymbol u} = \gamma W {\boldsymbol u} +
{\boldsymbol \epsilon} \label{eq:SEM}\\
{\boldsymbol y} &= X{\boldsymbol \beta} + (I-\gamma W)^{-1}{\boldsymbol \epsilon}\notag
\end{align}
If the likelihood ratio test rejects that the models are equivalent, then the SDM
should be used because its estimates are unbiased. One consequence of using the SDM
is that the estimated coefficients do not represent the

The researcher must assert $W$ *a priori*, though there are several
common specifications \citep{Dubin1998}. In this study we specify that tracts
with population-weighted centroids located within 1.8 Euclidean miles of each
other are considered neighbors, and the strength of the relationship is the
row-standardized (each row sums to 1) inverse of the distance between them:
\begin{equation}\label{eq:weightsrule}
W_{ij} = \begin{cases}
  {\frac{d_{ij}^{-1}} {\sum_{k=1}^n d_{ik}^{-1}}} & \text{for $d_{ij} \leq 1.8$ miles}\\
  0  & \text{for $d_{ij} > 1.8$ miles}
  \end{cases}
\end{equation}

```{r weights, warning = FALSE}
# This are developed in 03_logsums.Rmd
distances <- read_rds(here("data/distances.rds"))
W <- read_rds(here("data/W.rds"))
trMC <- trW(as(W, "CsparseMatrix"), type="MC")
```
This is a similar specification to that used by  \citet{tra_autoregression_paper}
in a study of home prices in Atlanta. We selected this specification after
comparing overall model likelihoods between it and a specification where tracts
sharing a boundary are considered neighbors.

@@

```{r termnames}
names_for_terms <- c(
  "(Intercept)", "log(Density)", "log(Income)",
  "Fulltime", "College-educated",  "Single Adults",
  "Youth (0-17)", "Young adults (18-34)",
  "Seniors (65+)", "Black population share",
  "Asian population share",
  "Hispanic population share", "Other Minorities",
  "$\\gamma$: spatial correlation","Physical activity",
  "Logsum", "Tweets", "Attributes", "10-min walk"
)

termnames <- tibble(
  name = names_for_terms,
  term = c("Intercept", "log(density)", "log(income)", "fulltime", "college",
           "single", "youth", "young_adults", "seniors", "black", "asian", 
           "hispanic", "other",
           "lag", "physact", "access_ls", "tweets_ls", "multi_ls", "walk_10TRUE"))
```

>>>>

We predict either the physical activity or the obesity rate ($y$) in a census
tract as a linear function of the tract's sociodemographic characteristics $X$
and accessibility to parks ($A$),
\begin{equation}\label{eq:themodel}
 y = X\beta + \beta_{a}A + \varepsilon
\end{equation}

As the observations are related to each other spatially, it is likely that this
process involves spatial spillovers of one kind or another. A complete treatment of
spatial econometrics is not warranted here; the reader is directed to
@LeSage2009 as well as @LeSage2014.
Suffice it to say that the presence of spatial data generating processes can
negatively affect econometric interpretation in a variety of ways. Each of these
process relies on spatial autoregression, where elements of some random
variable $x$ are spatially dependent on other elements, 
\begin{equation}\label{eq:autoregression}
 x = \rho W x + \varepsilon
\end{equation}
with the strength of dependence an estimated parameter $\rho$, the structure of 
dependence governed by the asserted matrix $W$ (details of $W$ follow below),
and some random uncorrelated residual $\varepsilon$. A collection of models is
available to represent spatial autodependence in the model residuals, the 
independent variables, and the dependent variable.

The specific spatial model applied is both an econometric choice
--- as improperly specified spatial models can lead to inconsistent estimates of
model parameters and standard errors --- as well as a philosophical choice about
the likely structure of spillovers in the problem at hand. In case of physical
activity and obesity rates, we believe that these outcomes
are locally determined: That is, a particular individual decides whether or not
to participate in physical activity independently of whether his or her neighbors
participate in physical activity. Mathematically, we assert that the autodependence
relationship on the dependent variable $y = \rho W y + \varepsilon$ has 
$\rho = 0$ in the current case.

With this relationship ruled out, it remains a possibility that the 
model residuals are spatially dependent,
\begin{equation}\label{eq:SEM}
 y = X\beta + \beta_a A + u; u = \lambda W u + \varepsilon
\end{equation}
that the outcome is partially dependent on the socioeconomic variables in *neighboring*
tracts,
\begin{equation}\label{eq:SLX}
 y = X\beta + \beta_a A + W X \lambda + \varepsilon
\end{equation}
or a linear combination of both.
\begin{equation}\label{eq:SDEM}
 y = X\beta + \beta_a A + W X \gamma + u; u = \lambda W u + \varepsilon 
\end{equation}
These models are referred as the spatial error model (SEM), the spatial lag of
$X$ model (SLX), and the spatial Durbin error model (SDEM). @LeSage2014 suggest
that a Hausman-style test of the estimates of $\beta$ derived from the OLS 
(Equation \ref{eq:themodel}) and SEM (Equation \ref{eq:SEM}) can identify whether
the estimates are consistent. If the $\beta$ estimates are dissimilar, both 
specifications are inappropriate as they contain untreated omitted variables.
Similarly, if the SLX and SDEM estimates are dissimilar it is an indication of 
missing variables in that specification. On the other hand if the estimates of
both pairs of models are similar and the residual correlation parameter $\lambda$
is non-zero, the model accounting for spatially dependent residuals will produce 
proper estimates of the model standard errors.

Note that in the lag-$X$ specifications (SLX and SDEM) we exclude the accessiblity 
component $A$, that is we do *not* consider that the accessibility
in a neighboring zone will have an effect on physical activity or obesity rates.
In a practical sense, this is because the accessibility term $A$ is itself spatially
determined.

### Spatial Weights

The spatial weights matrix $W$ is constructed of individual elements where
$w_{ij} = 0$ if observations $i$ and $j$ are spatially independent of each other,
and $w_{ij} > 0$ if $i$ and $j$ are spatially related in some way. 
@Dubin1998 presents details on constructing $W$, but in this study we
assume that tracts sharing at least one common border point are neighbors of
each other. The elements of $W$ are then row-standardized so that each observation
has an equal total influence from all its neighbors.

```{r weights, warning = FALSE}
# This are developed in 03_logsums.Rmd
distances <- read_rds(here("data/distances.rds"))
W <- read_rds(here("data/W.rds"))
trMC <- trW(as(W, "CsparseMatrix"), type="MC")
```

@@

```{r termnames}
base_termnames <- c(
  "(Intercept)", "log(Density)", "log(Income)", "Fulltime", "College-educated",  "Single Adults",
  "Youth (0-17)", "Young adults (18-34)", "Seniors (65+)", "Black population share",
  "Asian population share", "Hispanic population share", "Other Minorities"
)

names_for_terms <- c(
  base_termnames, str_c("$\\gamma$: ", base_termnames[-1]),
  "$\\lambda$: spatial correlation"
)

access_terms <- c("Logsum", "Tweets",  "Attributes", "10-min walk")
```
